Conversations in time: interactive visualisation to explore structured temporal data

### 2021-02-15 reviews/response.Rmd

---
title: Response to reviewers
author: RJournal 2020-171
output:
  UOARmdTemplates::memo:
    includes:
      in_header: preamble2.tex
fontfamily: mathpazo
---

We thank the Editor and reviewers for their comments which have improved this paper. The point by point description of changes are below: the reviewers' comments are in red and our response is in black.

# Reviewer 1 comments {-}

> Bottom of page 8: "Figure 3 peels the tidy module with" I'm not sure about the use of the word peels there?

We replaced "peels" with "reveals".

> Top of page 9: "The rigidity of tsibble, as the fundamental data infrastructure, warrants the validity of temporal data analysis for the later stage." Should "warrants" be replaced with "ensures" and perhaps replace "temporal data analysis for the later stage" with "subsequent temporal data analysis".

Done.

# Reviewer 2 comments {-}

> P4 L59 and the enhancements of -> and the enhancements by

Done.

> P6 L34. Second item missing full stop at the end.

Done.

> I think for Figure 4, you should mention that supplementary figure contains the gif if the animation does not work or perhaps include a link to a working animation.

Thanks. We added a sentence in Figure 4 caption "The animations are available with the supplementary materials online, and can also be viewed directly at <https://github.com/earowang/paper-tsibble/blob/master/img/animate-1.gif>."

> P15 L58. Mention `future_*` is from `furrr` package.

Done.

> P16 L24. Missing reference for `lubridate`.

Done.

### 2021-02-15 reviews/response.Rmd

---
title: Response to reviewers
author: RJournal 2020-171
output:
  UOARmdTemplates::memo:
    includes:
      in_header: preamble2.tex
fontfamily: mathpazo
---

We thank the reviewers for their comments which have improved this paper. The point by point description of changes are below: the reviewers' comments are in red and our response is in black.

@@

# Reviewer 1 comments {-}

> Please mention that tidyverts also contains the tsibbletalk package

The tidyverts only contains {tsibble}, {fable}, and {feasts}. The {tsibbletalk} packages is NOT part of the tidyverts suite.

> The sentence "The techniques implemented in work described in this paper utilise web technology." could be more specific. One fix could include moving the sentence in the Summary section "At the heart of the tsibbletalk package is a blending of the best bits from tsibble, crosstalk, plotly, and shiny." to an earlier section in the paper.

Thanks. Done.

> In the section "Linking between plots" you write "... linking is by time, ...". It seems to me that linking is by key, so all time series points associated to a particular key are linked.

Yes. We corrected from "by time" to "by key variables".

> The phrasing "to allow for structured specifications in the key" is unclear. It would be better to be more explicit saying that it is to define the crossing and/or nesting relationships among the key variables.

@@

Done.

> The phrasing "user-supplied session id" is wrong, a session means something very particular in shiny whereas the id argument if a module takes an id string (that should not be repeated in the same shiny namespace).

Thanks. We corrected "user-supplied session id" to "user-supplied id string".

@@

# Reviewer 2 comments {-}

> Within the article, I would like to see more discussion of the actual linking implementation. I believe the package is using the R6 class and creating a subclass of SharedData from 'crosstalk', but it would be nice to see this explained within the article itself. A little note on why this is the preferred solution would also be of interest to readers

@@

Done.

> The code snippets lack library calls or the package::function_in_package format, for reproducibility.

Thanks. Added missing {ggplot2}, {plotly}, {shiny}, and {feasts} library calls in the code snippets.

> In the "Overview of interactivity", no mention is made of the non-web-based interactivity made available by the 'grDevices' package (https://protect-au.mimecast.com/s/xMG5CGv0kQtJA4DNh7AMnU?domain=github.com).

We added the mention of {grDevices} in the "Overview of interactivity" section: "Systems without the web technology include grDevices ...".

### 2021-02-15 tsibbletalk.Rmd
started revising

Systems without the web technology include \CRANpkg{loon} [@R-loon], based on Tcl/Tk, and \pkg{cranvas} [@xie_reactive_2014] based on Qt. They offer a wide array of pre-defined interactions, such as selecting and zooming, to manipulate plots via mouse action, keyboard strokes, and menus. The \pkg{cranvastime} package [@cheng_enabling_2016] is an add-on to \pkg{cranvas}, which provides specialised interactions for temporal data, such as wrapping and mirroring.

The techniques implemented in work described in this paper utilise web technology.

@@

As opposed to one-to-one linking, \CRANpkg{tsibbletalk} defaults to categorical variable linking where selecting one or more observations in one category will broadcast to all other observations in this category. That is, linking is by time: within the time series plot, click on any data point, and  whole line will be highlighted in response. 
The `as_shared_tsibble()` uses `tsibble`'s `key` variables to achieve these types of linking.

@@

The `spec` argument in `as_shared_tsibble()` provides a means to construct hybrid linking, that incorporates hierarchical and categorical linking. A symbolic formula can be passed to the `spec` argument, to allow for structured specifications in the `key`. Adopting @Wilkinson1973's notation for factorial models, the `spec` follows the `/` and `*` operator conventions to declare nesting and crossing variables, respectively. The `spec` for the `aus_retail` data is therefore specified as `State * Industry` or `Industry * State`, which is the default for the presence of multiple `key` variables. If there is a hierarchy in the data, using `/` is required to indicate the parent-child relation, for a strictly one directional `parent/child`.

@@

```{r load-other-pkgs}
```

@@

This shiny module, decoupled to `tsibbleWrapUI()` and `tsibbleWrapServer()`, presents a clean interface and forms a reusable component that could be embedded in any shiny application. In general, a shiny module provides a vehicle for modularising shiny applications, relevant for both users and developers. As with all shiny modules, the first argument in both functions in \CRANpkg{tsibbletalk} requires a user-supplied session id that must be unique. The UI function `tsibbleWrapUI()` simply shows a slider that animates or controls the number of periods to be diced. The workhorse is the server function `tsibbleWrapServer()`, encapsulating the algorithm that transforms data and sends messages to update the plot accordingly. The `plot` argument expects a `ggplot` or `plotly` object, where one can plot data using either lines or other graphical elements (such as boxplots). As the function name suggests, a (shared) tsibble is needed to start the engine, so that the time `index` can be retrieved for dissection. The `period` option semantically takes a desired number of seasonal periods to be shifted, for example data shifted by "1 day", "2 days", or "1 week", etc. In other words, the `period` defines the grind level. For date-times (represented by `POSIXt`), the granularity ranges from fine "day" to a much coarser "year". The following code snippet generates Figure \ref{fig:wrap-ped}.

>>>>

Systems without the web technology include \CRANpkg{grDevices}, \CRANpkg{loon} [@R-loon], based on Tcl/Tk, and \pkg{cranvas} [@xie_reactive_2014] based on Qt. They offer a wide array of pre-defined interactions, such as selecting and zooming, to manipulate plots via mouse action, keyboard strokes, and menus. The \pkg{cranvastime} package [@cheng_enabling_2016] is an add-on to \pkg{cranvas}, which provides specialised interactions for temporal data, such as wrapping and mirroring.

The techniques implemented in work described in this paper utilise web technology, including \CRANpkg{crosstalk}, \CRANpkg{plotly}, and R \CRANpkg{shiny}.

@@

As opposed to one-to-one linking, \CRANpkg{tsibbletalk} defaults to categorical variable linking where selecting one or more observations in one category will broadcast to all other observations in this category. That is, linking is by key variables: within the time series plot, click on any data point, and  whole line will be highlighted in response. 
The `as_shared_tsibble()` uses `tsibble`'s `key` variables to achieve these types of linking.

@@

The `spec` argument in `as_shared_tsibble()` provides a means to construct hybrid linking, that incorporates hierarchical and categorical linking. A symbolic formula can be passed to the `spec` argument, to define the crossing and/or nesting relationships among the key variables. Adopting @Wilkinson1973's notation for factorial models, the `spec` follows the `/` and `*` operator conventions to declare nesting and crossing variables, respectively. The `spec` for the `aus_retail` data is therefore specified as `State * Industry` or `Industry * State`, which is the default for the presence of multiple `key` variables. If there is a hierarchy in the data, using `/` is required to indicate the parent-child relation, for a strictly one directional `parent/child`.

@@

@@

This shiny module, decoupled to `tsibbleWrapUI()` and `tsibbleWrapServer()`, presents a clean interface and forms a reusable component that could be embedded in any shiny application. In general, a shiny module provides a vehicle for modularising shiny applications, relevant for both users and developers. As with all shiny modules, the first argument in both functions in \CRANpkg{tsibbletalk} requires a user-supplied id string that must be unique. The UI function `tsibbleWrapUI()` simply shows a slider that animates or controls the number of periods to be diced. The workhorse is the server function `tsibbleWrapServer()`, encapsulating the algorithm that transforms data and sends messages to update the plot accordingly. The `plot` argument expects a `ggplot` or `plotly` object, where one can plot data using either lines or other graphical elements (such as boxplots). As the function name suggests, a (shared) tsibble is needed to start the engine, so that the time `index` can be retrieved for dissection. The `period` option semantically takes a desired number of seasonal periods to be shifted, for example data shifted by "1 day", "2 days", or "1 week", etc. In other words, the `period` defines the grind level. For date-times (represented by `POSIXt`), the granularity ranges from fine "day" to a much coarser "year". The following code snippet generates Figure \ref{fig:wrap-ped}.

### 2021-02-15 tsibbletalk.tex
started revising

Systems without the web technology include \CRANpkg{loon}
\citep{R-loon}, based on Tcl/Tk, and \pkg{cranvas}
\citep{xie_reactive_2014} based on Qt. They offer a wide array of
pre-defined interactions, such as selecting and zooming, to manipulate
plots via mouse action, keyboard strokes, and menus. The
\pkg{cranvastime} package \citep{cheng_enabling_2016} is an add-on to
\pkg{cranvas}, which provides specialised interactions for temporal
data, such as wrapping and mirroring.

The techniques implemented in work described in this paper utilise web
technology.

@@

As opposed to one-to-one linking, \CRANpkg{tsibbletalk} defaults to
categorical variable linking where selecting one or more observations in
one category will broadcast to all other observations in this category.
That is, linking is by time: within the time series plot, click on any
data point, and whole line will be highlighted in response. The
\texttt{as\_shared\_tsibble()} uses \texttt{tsibble}'s \texttt{key}
variables to achieve these types of linking.

@@

The \texttt{spec} argument in \texttt{as\_shared\_tsibble()} provides a
means to construct hybrid linking, that incorporates hierarchical and
categorical linking. A symbolic formula can be passed to the
\texttt{spec} argument, to allow for structured specifications in the
\texttt{key}. Adopting \citet{Wilkinson1973}'s notation for factorial
models, the \texttt{spec} follows the \texttt{/} and \texttt{*} operator
conventions to declare nesting and crossing variables, respectively. The
\texttt{spec} for the \texttt{aus\_retail} data is therefore specified
as \texttt{State\ *\ Industry} or \texttt{Industry\ *\ State}, which is
the default for the presence of multiple \texttt{key} variables. If
there is a hierarchy in the data, using \texttt{/} is required to
indicate the parent-child relation, for a strictly one directional
\texttt{parent/child}.

@@

\begin{Schunk}
\begin{Sinput}
p_tr <- tourism_shared %>%
  ggplot(aes(x = Month, y = Trips)) +
  geom_line(aes(group = Region), alpha = .5, size = .4) +
  facet_wrap(~ Purpose, scales = "free_y") +
  scale_x_yearmonth(date_breaks = "5 years", date_labels = "%Y")
\end{Sinput}
\end{Schunk}

@@

\begin{Schunk}
\begin{Sinput}
tourism_feat <- tourism_shared %>%
  features(Trips, feat_stl)
p_br <- tourism_feat %>%
  ggplot(aes(x = trend_strength, y = seasonal_strength_year)) +
  geom_point(aes(group = Region), alpha = .8, size = 2)
\end{Sinput}
\end{Schunk}

@@

\begin{Schunk}
\begin{Sinput}
subplot(p_l,
  subplot(
    ggplotly(p_tr, tooltip = "Region", width = 1100),
    ggplotly(p_br, tooltip = "Region", width = 1100),
    nrows = 2),
  widths = c(.4, .6)) %>%
  highlight(dynamic = TRUE)
\end{Sinput}
\end{Schunk}

@@

This shiny module, decoupled to \texttt{tsibbleWrapUI()} and
\texttt{tsibbleWrapServer()}, presents a clean interface and forms a
reusable component that could be embedded in any shiny application. In
general, a shiny module provides a vehicle for modularising shiny
applications, relevant for both users and developers. As with all shiny
modules, the first argument in both functions in \CRANpkg{tsibbletalk}
requires a user-supplied session id that must be unique. The UI function
\texttt{tsibbleWrapUI()} simply shows a slider that animates or controls
the number of periods to be diced. The workhorse is the server function
\texttt{tsibbleWrapServer()}, encapsulating the algorithm that
transforms data and sends messages to update the plot accordingly. The
\texttt{plot} argument expects a \texttt{ggplot} or \texttt{plotly}
object, where one can plot data using either lines or other graphical
elements (such as boxplots). As the function name suggests, a (shared)
tsibble is needed to start the engine, so that the time \texttt{index}
can be retrieved for dissection. The \texttt{period} option semantically
takes a desired number of seasonal periods to be shifted, for example
data shifted by ``1 day'', ``2 days'', or ``1 week'', etc. In other
words, the \texttt{period} defines the grind level. For date-times
(represented by \texttt{POSIXt}), the granularity ranges from fine
``day'' to a much coarser ``year''. The following code snippet generates
Figure \ref{fig:wrap-ped}.

\begin{Schunk}
\begin{Sinput}
p_line <- pedestrian20 %>%
  ggplot(aes(x = Date_Time, y = Count, colour = Lockdown)) +
  geom_line(size = .3) +
  facet_wrap(~ Sensor, scales = "free_y") +
  labs(x = "Date Time") +
  scale_colour_brewer(palette = "Dark2") +
  theme(legend.position = "none")

>>>>

Systems without the web technology include \CRANpkg{grDevices},
\CRANpkg{loon} \citep{R-loon}, based on Tcl/Tk, and \pkg{cranvas}
\citep{xie_reactive_2014} based on Qt. They offer a wide array of
pre-defined interactions, such as selecting and zooming, to manipulate
plots via mouse action, keyboard strokes, and menus. The
\pkg{cranvastime} package \citep{cheng_enabling_2016} is an add-on to
\pkg{cranvas}, which provides specialised interactions for temporal
data, such as wrapping and mirroring.

The techniques implemented in work described in this paper utilise web
technology, including \CRANpkg{crosstalk}, \CRANpkg{plotly}, and R
\CRANpkg{shiny}.

@@

As opposed to one-to-one linking, \CRANpkg{tsibbletalk} defaults to
categorical variable linking where selecting one or more observations in
one category will broadcast to all other observations in this category.
That is, linking is by key variables: within the time series plot, click
on any data point, and whole line will be highlighted in response. The
\texttt{as\_shared\_tsibble()} uses \texttt{tsibble}'s \texttt{key}
variables to achieve these types of linking.

@@

The \texttt{spec} argument in \texttt{as\_shared\_tsibble()} provides a
means to construct hybrid linking, that incorporates hierarchical and
categorical linking. A symbolic formula can be passed to the
\texttt{spec} argument, to define the crossing and/or nesting
relationships among the key variables. Adopting \citet{Wilkinson1973}'s
notation for factorial models, the \texttt{spec} follows the \texttt{/}
and \texttt{*} operator conventions to declare nesting and crossing
variables, respectively. The \texttt{spec} for the \texttt{aus\_retail}
data is therefore specified as \texttt{State\ *\ Industry} or
\texttt{Industry\ *\ State}, which is the default for the presence of
multiple \texttt{key} variables. If there is a hierarchy in the data,
using \texttt{/} is required to indicate the parent-child relation, for
a strictly one directional \texttt{parent/child}.

@@

\begin{Schunk}
\begin{Sinput}
library(ggplot2)
p_tr <- tourism_shared %>%
  ggplot(aes(x = Month, y = Trips)) +
  geom_line(aes(group = Region), alpha = .5, size = .4) +
  facet_wrap(~ Purpose, scales = "free_y") +
  scale_x_yearmonth(date_breaks = "5 years", date_labels = "%Y")
\end{Sinput}
\end{Schunk}

@@

\begin{Schunk}
\begin{Sinput}
library(feasts)
tourism_feat <- tourism_shared %>%
  features(Trips, feat_stl)
p_br <- tourism_feat %>%
  ggplot(aes(x = trend_strength, y = seasonal_strength_year)) +
  geom_point(aes(group = Region), alpha = .8, size = 2)
\end{Sinput}
\end{Schunk}

@@

\begin{Schunk}
\begin{Sinput}
library(plotly)
subplot(p_l,
  subplot(
    ggplotly(p_tr, tooltip = "Region", width = 1100),
    ggplotly(p_br, tooltip = "Region", width = 1100),
    nrows = 2),
  widths = c(.4, .6)) %>%
  highlight(dynamic = TRUE)
\end{Sinput}
\end{Schunk}

@@

This shiny module, decoupled to \texttt{tsibbleWrapUI()} and
\texttt{tsibbleWrapServer()}, presents a clean interface and forms a
reusable component that could be embedded in any shiny application. In
general, a shiny module provides a vehicle for modularising shiny
applications, relevant for both users and developers. As with all shiny
modules, the first argument in both functions in \CRANpkg{tsibbletalk}
requires a user-supplied id string that must be unique. The UI function
\texttt{tsibbleWrapUI()} simply shows a slider that animates or controls
the number of periods to be diced. The workhorse is the server function
\texttt{tsibbleWrapServer()}, encapsulating the algorithm that
transforms data and sends messages to update the plot accordingly. The
\texttt{plot} argument expects a \texttt{ggplot} or \texttt{plotly}
object, where one can plot data using either lines or other graphical
elements (such as boxplots). As the function name suggests, a (shared)
tsibble is needed to start the engine, so that the time \texttt{index}
can be retrieved for dissection. The \texttt{period} option semantically
takes a desired number of seasonal periods to be shifted, for example
data shifted by ``1 day'', ``2 days'', or ``1 week'', etc. In other
words, the \texttt{period} defines the grind level. For date-times
(represented by \texttt{POSIXt}), the granularity ranges from fine
``day'' to a much coarser ``year''. The following code snippet generates
Figure \ref{fig:wrap-ped}.

\begin{Schunk}
\begin{Sinput}
library(shiny)
p_line <- pedestrian20 %>%
  ggplot(aes(x = Date_Time, y = Count, colour = Lockdown)) +
  geom_line(size = .3) +
  facet_wrap(~ Sensor, scales = "free_y") +
  labs(x = "Date Time") +
  scale_colour_brewer(palette = "Dark2") +
  theme(legend.position = "none")
